Modal Verification Patterns for Systems

### 2025-06-16 paper.tex
+

\paragraph{In-Place Updates in Filesystems}In-place updating filesystems with journalling components have been specified and verified with an extended Hoare Logic to specify not only the pre/post conditions of the operations successful operations in the filesystem, but also the crash condition~\cite{chen-crash-16,chen2017verifying}, with later work extending this for concurrent filesystem implementations~\cite{chajed2021gojournal}. \textsf{Perennial}~\cite{chajed2019verifying} extends the weakest precondition of the \textsf{Iris} separation logic~\cite{jung2018iris,krebbers2017essence} to allow asserting facts about disk state at crashes (the \emph{crash condition}) and apply this extension to a journalling filesystem \textsf{GoJournal}~\cite{chajed2021gojournal}. Unlike \textsf{GoJournal}, \textsf{ModFs} relies on \textsf{CoW} semantics for consistency, eliminating the need for recovery logic.
Essentially, successfully writing the new superblock for a new snapshot is analagous to completing a valid journal entry, except it \emph{also} has the effect of updating the live filesystem data due to making the ``journal entries'' valid filesystem data and metadata blocks (an idea that goes back to early work on log-structured filesystems~\cite{Rosenblum1992LFS}).
In fact, we can think of the atomicity of the recovery from the crash state conceptually equal to the single block disk write persisting the change from $\textsf{Super}_0$ to $\textsf{Super}_1$ in Figure \ref{fig:cow_index_tree_update}, therefore, considering \textsc{CoW} semantics, we do not need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.
\paragraph{Consistency of Disk Layout} \todo{}

\paragraph{Memory Reclamation in Client Code}\todo{iso:speak about Amal's stuff}
\paragraph{Consistency of Reclaimed Memory}\todo{iso: speak about the operator for which the hybrid logic stuff is usefull }

\subsection{Speculative Observation in Systems Software}
\label{sec:speculativeobservation}
\section{Introduction}\label{sec:intro}
\todo[inline]{Colin says: These comments aren't the only places things need to change, but they're
specific places I have suggestions.
One thing to keep in mind is that the CfP lists two kinds of papers:
short papers are for "original research or case studies, maybe in the incubation phase,"
while long papers are for "original research, case studies, or surveys."
This is clearly in the survey category, so we (a) should take advantage of the extra space to
explain things clearly (15 pages, plus 2 pages for bibliography), and (b) lean into the survey nature explicitly.
}

\todo[inline]{I think one of the biggest things we can do to reframe this is to explain the broader vision
here before jumping into contributions.
This is an audience that knows dynamic logic and modal logic more broadly, but perhaps not everyone is familiar
with hybrid logic, and many will not be familiar with systems concepts.
This is the place to provide high-level intuition about the domain we're surveying, tailored for this audience.
I'd talk about dynamic logic in general, but then highlight the need for reasoning specific to certain
compute resources to produce manageable specifications and proofs.
Think big picture: why is this space interesting to logicians?
}

@@

\subsection{Speculation in Consistency}
%\subsection{Sharing} Sharing is one of the fundamental reasons for kernels to {virtualize} resources,
%for creating the illusion of exclusive access to virtualized versions of shared resources, including CPU time (via preemption).
%An implementation of sharing relies on two concepts: \emph{indirection} and \emph{coordination}. Indirection is a method that allows the mapping of resources of one type into another. By doing so, it allows treating the limited resources (e.g. hardware resources such as physical memory) as if there are more than they actually are (e.g. via virtual memory addresses and paging). Consequently, virtualization for the sake of ensuring availability requires \emph{sharing} of the resources that exist in reality.
%%\paragraph{Isolation Hand-in-Hand with Sharing}
%%\Ismail{put a few words on abstractions such as address space, stack region, snapshots ,etc.}
%%\Colin{Not sure what you want here?}

>>>>

\begin{figure}
    \includegraphics[width=\columnwidth]{journalling.png}
    \caption{A Filesystem Utilizing \emph{Journalling} for Consistent Disk-Layout ~\cite{chajed2021gojournal}}
    \label{fig:journal}
\end{figure}

\paragraph{Different Models in Ensuring Disk Consistency of Filesystems}In-place updating filesystems with journalling components have been specified and verified with an extended Hoare Logic to specify not only the pre/post conditions of the operations successful operations in the filesystem, but also the crash condition~\cite{chen-crash-16,chen2017verifying}, with later work extending this for concurrent filesystem implementations~\cite{chajed2021gojournal}. \textsf{Perennial}~\cite{chajed2019verifying} extends the weakest precondition of the \textsf{Iris} separation logic~\cite{jung2018iris,krebbers2017essence} to allow asserting facts about disk state at crashes (the \emph{crash condition}) and apply this extension to a journalling filesystem \textsf{GoJournal}~\cite{chajed2021gojournal}. Unlike \textsf{GoJournal}, \textsf{ModFs} relies on \textsf{CoW} semantics for consistency, eliminating the need for recovery logic.
Essentially, successfully writing the new superblock for a new snapshot is analagous to completing a valid journal entry, except it \emph{also} has the effect of updating the live filesystem data due to making the ``journal entries'' valid filesystem data and metadata blocks (an idea that goes back to early work on log-structured filesystems~\cite{Rosenblum1992LFS}).
In fact, we can think of the atomicity of the recovery from the crash state conceptually equal to the single block disk write persisting the change from $\textsf{Super}_0$ to $\textsf{Super}_1$ in Figure \ref{fig:cow_index_tree_update}, therefore, considering \textsc{CoW} semantics, we do not need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.
\paragraph{Memory Reclamation in Client Code} Allocating memory locations for a client code leaves the 
\paragraph{Consistency of Reclaimed Memory}\todo{iso: speak about the operator for which the hybrid logic stuff is usefull }

@@

\subsection{Speculation in Consistency} \todo{speak about the weak-mem modalities in here.}
%\subsection{Sharing} Sharing is one of the fundamental reasons for kernels to {virtualize} resources,
%for creating the illusion of exclusive access to virtualized versions of shared resources, including CPU time (via preemption).
%An implementation of sharing relies on two concepts: \emph{indirection} and \emph{coordination}. Indirection is a method that allows the mapping of resources of one type into another. By doing so, it allows treating the limited resources (e.g. hardware resources such as physical memory) as if there are more than they actually are (e.g. via virtual memory addresses and paging). Consequently, virtualization for the sake of ensuring availability requires \emph{sharing} of the resources that exist in reality.
%%\paragraph{Isolation Hand-in-Hand with Sharing}
%%\Ismail{put a few words on abstractions such as address space, stack region, snapshots ,etc.}
%%\Colin{Not sure what you want here?}
